Erratum: ``Supermassive black holes and their host spheroids \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram'' (2016, ApJ, 817, 21)

### 2016-03-11 papers/MbhMsph/erratum/mm.tex
erratum

\documentclass[preprint2]{emulateapj}

\usepackage{natbib}
\bibliographystyle{apj}
\usepackage{longtable}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{color}
\usepackage{hyperref}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{gsavorgn@astro.swin.edu.au}
\newcommand{\fitfigurewidth}{0.8\textwidth}

\shorttitle{Red and blue $M_{\rm BH} - M_{\rm *,sph}$ sequence}

\shortauthors{Savorgnan et al.}

@@

\title{Supermassive black holes and their host spheroids \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram}

@@

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, $M_{\rm BH}$, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between $M_{\rm BH}$ 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
While most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies,
our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$  
and define a tight 
\emph{red sequence} with intrinsic scatter $\epsilon_{(M_{\rm BH}|M_{\rm *,sph})} = 0.43 \pm 0.06\rm~dex$ 
and a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.68 \pm 0.04\%$, 
i.e.~a $\pm 2\sigma$ range of 0.1--5\%.
At the low-mass end, the bulges of late-type galaxies define a much steeper 
\emph{blue sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$ and $M_{\rm BH}/M_{\rm *,sph}$ equal to $0.02\%$ at $M_{\rm BH} \approx 10^6~\rm M_\odot$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) our S\'ersic galaxy sample follows $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, 
a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by the current bulge sample with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies 
-- once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

\end{abstract}

\keywords{black hole physics; galaxies: bulges; galaxies: elliptical and lenticular, cD; galaxies: evolution; galaxies: structure}

\section{Introduction}
\label{sec:int}
A quarter of a century ago, 
\cite{dressler1989} foresaw a ``rough scaling of black hole mass with the mass of the spheroidal component'' of galaxies, 
as suggested by the sequence of five galaxies (M87, M104, M31, M32 and the Milky Way). 
\cite{yee1992} then announced a linear relation between what was effectively black hole mass and galaxy mass for high-luminosity, bulge-dominated early-type galaxies 
radiating near the Eddington limit.
This ``rough scaling'' was a premature version of the early correlations between black hole mass, $M_{\rm BH}$,  
and host spheroid luminosity, $L_{\rm sph}$, and also host spheroid mass, $M_{\rm sph}$ 
\citep{kormendyrichstone1995,magorrian1998,marconihunt2003,haringrix2004}. 
These initial studies were dominated by high-mass, early-type galaxies, 
for which they too reported a quasi-linear $M_{\rm BH} - M_{\rm sph}$ relation. 
Subsequent studies of the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm sph}$ diagrams 
(\citealt{ferrareseford2005,lauer2007,graham2007,gultelkin2009,sani2011,beifiori2012,erwingadotti2012,
vika2012,vandenbosch2012,mcconnellma2013,kormendyho2013})
continued to use galaxy samples dominated by high-mass, early-type systems with $M_{\rm BH} \gtrsim 0.5 \times 10^8~\rm M_\odot$, 
and they too recovered a near-linear relation. 
However, the consensus about a linear $M_{\rm BH} - M_{\rm sph}$ correlation was not unanimous. 
Some studies had reported a slope steeper than one,  
or noticed that the low-mass spheroids were offset to the right of (or below) the relation traced by the high-mass spheroids 
\citep{laor1998,wandel1999,laor2001,ryan2007}.
\cite{graham2012bent}, \cite{grahamscott2013} and \cite{scott2013} found two distinct trends in the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm sph}$ diagrams:  
a linear and a super-quadratic correlation at the high- and low-mass end, respectively\footnote{Readers 
interested in an extensive review about the early discovery and successive improvements of these correlations 
should consult \cite{graham2015bulges}.}. \\
Recently, \cite{lasker2014data,lasker2014anal} derived $2.2~\rm \mu m$ bulge luminosities for 35 galaxies 
(among which only 4 were classified as spiral galaxies), 
and reported a slope below unity for their $M_{\rm BH} - L_{\rm sph}$ relation. 
They also claimed that the black hole mass correlates equally well with the total galaxy luminosity 
as it does with the bulge luminosity.  \\
The $M_{\rm BH} - L_{\rm sph}$ relation for early-type (elliptical + lenticular) galaxies can be predicted by combining two other correlations that involve 
the bulge stellar velocity dispersion, $\sigma$.
One of these is the $M_{\rm BH} - \sigma$ relation \citep{ferraresemerritt2000,gebhardt2000},
which can be described with a single power-law ($M_{\rm BH} \propto \sigma^{5-6}$) 
over a wide range in velocity dispersion ($70-350~\rm km~s^{-1}$, e.g.~\citealt{graham2011,mcconnell2011,grahamscott2013}).
The other is the $L_{\rm sph} - \sigma$ relation, 
which has long been known to be a ``double power-law'', 
with $L_{\rm sph} \propto \sigma^{5-6}$ at the luminous end\footnote{Recent 
work has the $M_{\rm BH} - \sigma$ correlation as steep as $M_{\rm BH} \propto \sigma^{6.5}$ \citep{savorgnangraham2015},
and the high-luminosity end of the $L_{\rm sph} - \sigma$ correlation as steep as $L_{\rm sph} \propto \sigma^{8}$ \citep{monterodorta2015}.} 
\citep{schechter1980,malumuthkrishner1981,vonderlinden2007,lauer2007lumell,liu2008}
and $L_{\rm sph} \propto \sigma^2$ at intermediate and faint luminosities 
\citep{davies1983,held1992,matkovicguzman2005,derijcke2005,balcells2007screl,chilingarian2008,forbes2008,cody2009,tortora2009,kourkchi2012}. 
The change in slope of the $L_{\rm sph} - \sigma$ relation occurs at $M_B \approx -20.5\rm~mag$, 
corresponding to $\sigma \approx 200~\rm km~s^{-1}$. 
The $M_{\rm BH} - L_{\rm sph}$ relation should therefore be better described by a ``broken'', rather than a single power-law: 
with $M_{\rm BH} \propto L_{\rm sph}^{2.5}$ at the low-luminosity end, 
and $M_{\rm BH} \propto L_{\rm sph}^1$ at the high-luminosity end.  
Due to the scatter in the $M_{\rm BH} - L_{\rm sph}$ (or $M_{\rm BH} - M_{\rm sph}$) diagram, 
studies that have not sufficiently probed below $M_{\rm BH} \approx 10^7\rm~M_\odot$ 
can easily miss the change in slope occuring at $M_{\rm BH} \approx 10^{(8 \pm 1)}\rm~M_\odot$, 
and erroneously recover a single log-linear relation. \\
When \cite{graham2012bent} pointed out this overlooked inconsistency between these linear and bent relations, 
he identified two different populations of galaxies, 
namely the core-S\'ersic spheroids \citep{graham2003coresersicmodel,trujillo2004coresersicmodel} and the S\'ersic 
spheroids\footnote{Core-S\'ersic spheroids have partially depleted cores relative to their outer S\'ersic light profile, 
whereas S\'ersic spheroids have no central deficit of stars.},
and attributed the change in slope (from super-quadratic to linear) to their different formation mechanisms. 
In this scenario, core-S\'ersic spheroids are built in dry merger events 
where the black hole and the bulge grow at the same pace, increasing their mass in lock steps ($M_{\rm BH} \propto L_{\rm sph}^1$), 
whereas S\'ersic spheroids originate from gas-rich processes 
in which the mass of the black hole increases more rapidly than the mass of its host spheroid ($M_{\rm BH} \propto L_{\rm sph}^{2.5}$). \\
\citeauthor{grahamscott2013} (\citeyear{grahamscott2013}, hereafter GS13) and \citeauthor{scott2013} (\citeyear{scott2013}) 
presented separate power-law linear regressions 
for the S\'ersic and core-S\'ersic spheroids in the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm *,sph}$ 
(spheroid stellar mass) diagrams, probing down to $M_{\rm BH} \approx 10^6\rm~M_\odot$. 
To obtain their dust-corrected \emph{bulge} magnitudes, they did not perform bulge/disc decompositions, 
but converted the $B-$band and $K_S-$band observed, total \emph{galaxy} magnitudes into bulge magnitudes 
using a mean statistical bulge-to-total ratio based on each object's morphological type and disc 
inclination\footnote{While this resulted in individual bulge magnitudes not being exactly correct, 
their large sample size allowed them to obtain a reasonable $M_{\rm BH} - L_{\rm sph}$ relation for the ensemble.}. 
These mean statistical bulge-to-total ratios were obtained from the results of two-component (S\'ersic-bulge/exponential-disk) decompositions in the literature. \\
Here we investigate in more detail 
the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm *,sph}$ diagrams 
using state-of-the-art galaxy decompositions (Savorgnan \& Graham 2015, hereafter \emph{Paper I}) 
for galaxies with directly measured black hole masses.
Our galaxies are large and nearby, which allows us to perform accurate multicomponent decompositions 
(instead of simple bulge/disk decompositions). 
Our decompositions were performed on $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery, 
which is an excellent proxy for the stellar mass, superior to the $K-$band (\citealt{sheth2010}, and references therein).
Nine of our galaxies have $M_{\rm BH} \lesssim 10^7\rm~M_\odot$, 
which allows us to better constrain the slope of the correlation at the low-mass end.
Furthermore, our galaxy sample includes 17 spiral galaxies, 
representing a notable improvement over past studies dominated by early-type galaxies. 
In a forthcoming paper, we will explore the relation between the black hole mass and the bulge dynamical mass, 
$M_{\rm dyn,sph} \propto R_{\rm e} \sigma^2$, and address the issue of a black hole fundamental plane.

\section{Data}
\label{sec:data}
Our galaxy sample (see Table \ref{tab:sample}) 
consists of 66 objects for which a dynamical measurement of the black hole mass had been tabulated in the literature 
(by GS13 or \citealt{rusli2013bhmassesDM}) at the time that we started this project, 
and for which we were able to obtain useful spheroid parameters from $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery. \\
Spheroid magnitudes were derived from our state-of-the-art galaxy decompositions, which take into account 
bulge, disks, spiral arms, bars, rings, halo, extended or unresolved nuclear source and partially depleted core. 
Kinematical information \citep{atlas3dIII,scott2014,arnold2014} was used 
to confirm the presence of rotationally supported disk components in most early-type (elliptical + lenticular) galaxies, 
and to identify their extent 
(intermediate-scale disks that are fully embedded in the bulge, 
or large-scale disks that encase the bulge and dominate the light at large radii). 
It is worth stressing that, contrary to common knowledge, the majority of ``elliptical'' galaxies contain disks,
i.e.~they are not single-component spheroidal systems.
\emph{Paper I} presents the dataset used here, 
including details about the data reduction process and the galaxy modelling technique that we developed. 
It also discusses how we estimated the uncertainties\footnote{By comparing, for each of our galaxies, the measurements of the bulge magnitude 
obtained by different authors with that obtained by us, we estimated the uncertainties on the bulge magnitudes, 
in effect taking into account systematic errors. 
Systematic errors include incorrect sky subtraction, inaccurate masking of contaminating sources, imprecise description of the PSF, 
erroneous choice of model components (for example, when failing to identify a galaxy subcomponent and thus omitting it in the model, 
or when describing a galaxy sub-component with an inadequate function), 
the radial extent of the surface brightness profile and one's sampling of this. 
Most of these factors are not included in popular 2D fitting codes which report only the statistical errors associated with their fitted parameters. 
In fact, when performing multi-component decomposition of high signal-to-noise images of nearby -- therefore well spatially resolved -- galaxies, 
errors are dominated by systematics rather than Poisson noise. 
Unlike many papers, we believe that we have not under-estimated the uncertainties associated to the bulge best-fit parameters. } 
on the bulge magnitudes, and presents the individual 66 galaxy decompositions, 
along with a comparison and discussion of past decompositions. \\
Bulge luminosities\footnote{Following \cite{sani2011}, absolute luminosities were calculated 
assuming a $3.6\rm~\mu m$ solar absolute magnitude of $3.25\rm~mag$. 
Absolute luminosities were not corrected for cosmological redshift dimming 
(this correction would be as small as $-0.02\rm~mag$ for galaxies at a distance of $40\rm~Mpc$ 
or $-0.05\rm~mag$ for galaxies at a distance of $100\rm~Mpc$).} 
(Table \ref{tab:sample}) from \emph{Paper I} were converted into stellar masses 
using a constant $3.6\rm~\mu m$ mass-to-light ratio, $\Gamma_{3.6} = 0.6$ \citep{meidt2014}.
We additionally explored a more sophisticated way to compute mass-to-light ratios, 
using the color-$\Gamma_{3.6}$ relation published by 
\citeauthor{meidt2014} (\citeyear{meidt2014}, their equation 4), 
which allows one to estimate $\Gamma_{3.6}$ of a galaxy from its $[3.6] - [4.5]$ color. 
Individual $[3.6] - [4.5]$ colors\footnote{These are integrated $[3.6] - [4.5]$ colors, measured in a circular aperture 
within each galaxy's effective radius.} were taken from 
\citeauthor{peletier2012} (\citeyear{peletier2012}, column 8 of their Table 1) 
when available for our galaxies, 
or were estimated from the bulge stellar velocity dispersion, $\sigma$, 
using the color-$\sigma$ relation presented by \citeauthor{peletier2012} (\citeyear{peletier2012}, their Figure 6).
We found that the range in $[3.6] - [4.5]$ color is small ($0.06\rm~mag$), 
and thus the range in $\Gamma_{3.6}$ is also small ($0.04$).
After checking that using a single $\Gamma_{3.6} = 0.6$, independent of $[3.6] - [4.5]$ color, 
does not significantly affect the results of our analysis, 
we decided to use individual, color-dependent mass-to-light ratios. \\
For each galaxy, the total luminosity (or galaxy luminosity, $L_{\rm gal}$) is the sum of the luminosities of all its sub-components. 
Due to the complexity of their modelling, 
four galaxies (see Table \ref{tab:sample}, column 7) had their galaxy luminosities 
underestimated\footnote{These four cases are discussed in \emph{Paper I}.}, 
which are given here as lower limits. 
Following GS13, we assumed a fixed uncertainty ($0.25\rm~mag$) for the absolute galaxy magnitude $MAG_{\rm gal}$. \\
The morphological classification (E = elliptical; E/S0 = elliptical/lenticular; S0 = lenticular; S0/Sp = lenticular/spiral; Sp = spiral; and ``merger'') 
follows from the galaxy models presented in \emph{Paper I}. 
Throughout this paper we will refer to early-type galaxies (E+S0) and late-type galaxies (Sp). 
Two galaxies classified as E/S0 are obviously included in the early-type bin, 
whereas two galaxies classified as S0/Sp and another two classified as mergers are included in neither the early- nor the late-type bin.\\
The S\'ersic/core-S\'ersic classification presented in this work 
comes from the compilation of \citet{savorgnangraham2015},
who identified partially depleted cores according to the same criteria used by GS13.
When no high-resolution image analysis was available from the literature, 
they inferred the presence of a partially depleted core based on the stellar velocity dispersion:
a spheroid is classified as core-S\'ersic if $\sigma > 270\rm~km~s^{-1}$,
or as S\'ersic if $\sigma < 166\rm~km~s^{-1}$. 
All of the galaxies with velocity dispersions between these two limits had high-resolution images available.

\input{table_sample.tex}

\section{Analysis}
\label{sec:anal}
We performed a linear regression analysis of the $M_{\rm BH} - L_{\rm gal}$ (see Table \ref{tab:lreggal}), 
$M_{\rm BH} - L_{\rm sph}$ (see Table \ref{tab:lregsph}) and $M_{\rm BH} - M_{\rm *,sph}$ (see Table \ref{tab:lregmass}) data,
using the BCES code from \cite{akritasbershady1996}. 
We also repeated the analysis using both the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}. 
All of these three linear regression routines account for the intrinsic scatter, 
but only the last two allow one to quantify it.
We report linear regressions, both symmetrical and non-symmetrical, 
for S\'ersic/core-S\'ersic and for early/late-type galaxies.
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas non-symmetrical forward ($M_{\rm BH}|X$) regressions -- 
which minimize the scatter in the $\log(M_{\rm BH})$ direction -- 
are best used to predict black hole masses.

\section{Results and discussion}
\label{sec:res}

\subsection{Black hole mass -- galaxy luminosity}
The $M_{\rm BH} - L_{\rm gal}$ diagram is shown in Figure \ref{fig:mbhmaggal}.
Four spiral galaxies had their total luminosities underestimated (see Table \ref{tab:sample}) 
and thus are not included in the linear regression analysis (see Table \ref{tab:lreggal}). \\
\cite{lasker2014anal} analyzed a sample of 35 galaxies, among which only four were classified as spiral galaxies, 
and claimed that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations, 
which they fit with a single power-law, have consistent intrinsic scatter.
Here, instead, thanks to our galaxy sample that includes 17 spiral galaxies, 
we show that the claim made by \cite{lasker2014anal} is valid only for early-type galaxies. 
That is, when considering only early-type galaxies, 
we find that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations have the same level of intrinsic scatter. 
However, our $M_{\rm BH} - L_{\rm sph}$ relation for all 66 galaxies, irrespective of their morphological type, 
has an intrinsic scatter $\epsilon_{(Y|X)} = 0.51 \pm 0.06\rm~dex$ (forward linear regression) 
and $\epsilon_{(X|Y)} = 0.60 \pm 0.09\rm~dex$ (inverse linear regression), 
whereas our $M_{\rm BH} - L_{\rm gal}$ relation for 62 ($=66-4$) galaxies has $\epsilon_{(Y|X)} = 0.63 \pm 0.07\rm~dex$ 
and $\epsilon_{(X|Y)} = 0.91 \pm 0.17\rm~dex$.
Because the value of the intrinsic scatter depends on the size of the uncertainties associated with the absolute magnitudes\footnote{The 
smaller (larger) the uncertainties, the larger (smaller) the intrinsic scatter.}, 
we tested the robustness of our conclusion by increasing the uncertainties associated with the galaxy absolute magnitudes\footnote{The 
value of the intrinsic scatter obviously depends also on the size of the uncertainties associated with the black hole masses. 
However, black hole masses and their uncertainties have been estimated by various authors using different methods, 
thus we have limited to no control on their values.}  
(we originally assumed $0.25\rm~mag$).
The intrinsic scatter of the $M_{\rm BH} - L_{\rm gal}$ relation only becomes smaller than that of the $M_{\rm BH} - L_{\rm sph}$ relation 
when assuming an uncertainty larger than $0.7\rm~mag$ for $L_{\rm gal}$, 
which would be significantly larger than the typical value commonly recognized in the literature, 
and would also oddly exceed the typical uncertainty that we estimated for $L_{\rm sph}$. 
Hence, we conclude that our determination of the relative intrinsic scatter is reliable, 
and that $M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ only for early-type galaxies\footnote{The majority 
of our early-type galaxies are elliptical galaxies, some of which have a bulge-to-total ratio close to 1 ($L_{\rm gal} \simeq L_{\rm sph}$).
One might wonder whether this constitutes a bias in our analysis.
However, we checked that $M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ 
not only for early-type (elliptical + lenticular) galaxies, 
but also for lenticular galaxies only. }, 
but not for all (early+late-type) galaxies.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{mbh_vs_mag_tot.pdf}
\caption{Black hole mass plotted against $3.6\rm~\mu m$ galaxy absolute magnitude. 
Symbols are coded according to the galaxy morphological type: red circle = E, red star = E/S0, 
red upward triangle = S0, blue downward triangle = S0/Sp, blue square = Sp, black diamond = merger. 
Empty symbols represent core-S\'ersic spheroids, whereas filled symbols are used for S\'ersic spheroids. 
Four spiral galaxies had their magnitudes overestimated (luminosities underestimated) and are shown as upper limits. 
The red dashed line indicates the BCES bisector linear regression for the 45 early-type galaxies (E+S0), 
with the red shaded area denoting its $1\sigma$ uncertainty. 
$M_{\rm BH}$ correlates equally well with $L_{\rm gal}$ and $L_{\rm sph}$ only for early-type galaxies, 
but not for all (early+late-type) galaxies. }
\label{fig:mbhmaggal}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{mbh_vs_mag_sph.pdf}
\caption{Black hole mass plotted against $3.6\rm~\mu m$ spheroid absolute magnitude. 
Symbols have the same meaning as in Figure \ref{fig:mbhmaggal}.
The red dashed line indicates the BCES bisector linear regression for the spheroidal component of the 45 early-type (E+S0) galaxies, 
with the red shaded area denoting its $1\sigma$ uncertainty. 
The blue solid line shows the BCES bisector linear regression for the bulges of the 17 late-type (Sp) galaxies, 
with the blue shaded area denoting its $1\sigma$ uncertainty. 
The black dashed-dotted and dotted lines represent the BCES bisector linear regressions for the core-S\'ersic and S\'ersic spheroids, respectively.}
\label{fig:mbhmagsph}
\end{center}
\end{figure}

\begin{table*}
\centering
\caption{Linear regression analysis of the $M_{\rm BH} - L_{\rm gal}$ diagram.}
\begin{tabular}{llccccc}
\tableline\tableline
{\bf Subsample (size)} & {\bf Regression} & $\boldsymbol \alpha$ & $\boldsymbol \beta$ & $\boldsymbol \langle MAG_{\rm gal} \rangle$ & $\boldsymbol \epsilon$ & $\boldsymbol \Delta$ \\ 
\tableline 
\\
  & \multicolumn{6}{l}{$\log[M_{\rm BH}/{\rm M_\odot}] = \alpha + \beta[(MAG_{\rm gal} - \langle MAG_{\rm gal} \rangle)/{\rm mag}]$} \\ [0.5em]
 All (62)               & BCES $(Y|X)$   & $8.26 \pm 0.08$ & $-0.49 \pm 0.06$ & $-24.78$ & $-$ & $0.64$ \\
                        & mFITEXY $(Y|X)$  & $8.26^{+0.08}_{-0.08}$ & $-0.49^{+0.06}_{-0.07}$ & $-24.78$ & $0.61^{+0.07}_{-0.06}$ & $0.64$ \\
                        & {\tt linmix\_err} $(Y|X)$  & $8.26 \pm 0.09$ & $-0.49 \pm 0.07$ & $-24.78$ & $0.63 \pm 0.07$ & $0.64$ \\ [0.5em]
                        & BCES $(X|Y)$   & $8.26 \pm 0.12$ & $-1.01 \pm 0.15$ & $-24.78$ & $-$ & $0.92$ \\
                        & mFITEXY $(X|Y)$  & $8.26^{+0.11}_{-0.12}$ & $-1.03^{+0.13}_{-0.16}$ & $-24.78$ & $0.88^{+0.10}_{-0.08}$ & $0.93$ \\
                        & {\tt linmix\_err} $(X|Y)$  & $8.26 \pm 0.12$ & $-1.02 \pm 0.15$ & $-24.78$ & $0.91 \pm 0.17$ & $0.93$ \\ [0.5em]
                        & BCES Bisector  & $8.26 \pm 0.09$ & $-0.72 \pm 0.07$ & $-24.78$ & $-$ & $0.71$ \\
                        & mFITEXY Bisector & $8.26^{+0.10}_{-0.10}$ & $-0.73^{+0.09}_{-0.10}$ & $-24.78$ & $-$    & $0.71$ \\
                        & {\tt linmix\_err} Bisector & $8.26 \pm 0.10$ & $-0.72 \pm 0.07$ & $-24.78$ & $-$    & $0.71$ \\ [0.5em]

 Early-type (E+S0) (45) & BCES $(Y|X)$   & $8.56 \pm 0.07$ & $-0.44 \pm 0.05$ & $-24.88$ & $-$ & $0.45$ \\
                        & mFITEXY $(Y|X)$  & $8.56^{+0.06}_{-0.06}$ & $-0.42^{+0.05}_{-0.05}$ & $-24.88$ & $0.41^{+0.06}_{-0.05}$ & $0.45$ \\
                        & {\tt linmix\_err} $(Y|X)$  & $8.56 \pm 0.07$ & $-0.42 \pm 0.06$ & $-24.88$ & $0.43 \pm 0.06$ & $0.45$ \\ [0.5em]
                        & BCES $(X|Y)$   & $8.56 \pm 0.08$ & $-0.64 \pm 0.05$ & $-24.88$ & $-$ & $0.53$ \\
                        & mFITEXY $(X|Y)$  & $8.56^{+0.08}_{-0.08}$ & $-0.66^{+0.07}_{-0.08}$ & $-24.88$ & $0.51^{+0.07}_{-0.06}$ & $0.55$ \\
                        & {\tt linmix\_err} $(X|Y)$  & $8.56 \pm 0.09$ & $-0.65 \pm 0.08$ & $-24.88$ & $0.53 \pm 0.10$ & $0.54$ \\ [0.5em]
                        & BCES Bisector  & $8.56 \pm 0.07$ & $-0.53 \pm 0.04$ & $-24.88$ & $-$ & $0.47$ \\
                        & mFITEXY Bisector & $8.56^{+0.07}_{-0.07}$ & $-0.54^{+0.06}_{-0.06}$ & $-24.88$ & $-$    & $0.47$ \\
                        & {\tt linmix\_err} Bisector & $8.56 \pm 0.08$ & $-0.53 \pm 0.05$ & $-24.88$ & $-$    & $0.47$ \\ [0.5em]

\tableline 
\tableline
\end{tabular}
\label{tab:lreggal} 
\tablecomments{For each subsample, we indicate $\langle MAG_{\rm gal} \rangle$, its average value of galaxy magnitudes. 
In the last two columns, we report $\epsilon$, the intrinsic scatter, and $\Delta$, the total rms scatter in the $\log(M_{\rm BH})$ direction. 
Four spiral galaxies had their luminosities underestimated and thus are not included in the linear regression analysis 
(the sample of all galaxies contains 66-4=62 objects). 
When considering all galaxies, irrespective of their morphological type, 
the $M_{\rm BH} - L_{\rm gal}$ correlation is weaker than the $M_{\rm BH} - L_{\rm sph}$ correlation, in terms of intrinsic scatter. 
However, when considering only early-type galaxies, the $M_{\rm BH} - L_{\rm gal}$ and $M_{\rm BH} - L_{\rm sph}$ correlations 
have consistent intrinsic scatter. }
\end{table*}

\subsection{Black hole mass -- spheroid luminosity}
The $M_{\rm BH} - L_{\rm sph}$ diagram is shown in Figure \ref{fig:mbhmagsph}, 
and the linear regression analysis is presented in Table \ref{tab:lregsph}. 
S\'ersic and core-S\'ersic spheroids have slopes consistent with each other (within their $1\sigma$ uncertainties), 
in disagreement with the findings of GS13. 
The slope that we obtained for core-S\'ersic spheroids ($M_{\rm BH} \propto L_{\rm sph}^{1.18 \pm 0.20}$) 
is consistent with the slope reported by GS13 in the $K_s$-band for the same population ($M_{\rm BH} \propto L_{\rm sph}^{1.10 \pm 0.20}$). 
However, the slope that we determined for S\'ersic spheroids ($M_{\rm BH} \propto L_{\rm sph}^{1.53 \pm 0.20}$) 
is notably shallower than that found by GS13 ($M_{\rm BH} \propto L_{\rm sph}^{2.73 \pm 0.55}$). \\
Although the S\'ersic/core-S\'ersic classification used by GS13 slightly differs\footnote{The classification has changed for the galaxies 
NGC 1316, NGC 1332 and NGC 3998.} from the classification used here, 
the main cause of the inconsistency is that the bulge-to-total ratios obtained from our galaxy decompositions 
are different from those assumed by GS13 to convert galaxy luminosities into bulge luminosities.
Our bulge-to-total ratios for low-luminosity S\'ersic spheroids ($3.6\rm~\mu m$ $MAG_{\rm sph} \gtrsim -22 \rm~mag$) 
are smaller than those used by GS13. 
The host galaxies of such bulges are late-type, spiral galaxies, 
which typically present a complex morphology (bars, double bars, embedded disks, nuclear components, etc).
Our galaxy models account for the extra components, 
while the average bulge-to-total ratios of GS13 were based on less sophisticated S\'ersic-bulge/exponential-disk decompositions 
which overestimated the bulge luminosity.
This results in our bulge magnitudes being on average $\sim$$0.7\rm~mag$ fainter than in GS13, after accounting for the different wavelength of the data.
At the same time, our bulge-to-total ratios for the high-luminosity S\'ersic spheroids ($3.6\rm~\mu m$ $MAG_{\rm sph} \lesssim -24 \rm~mag$) 
are on average larger than those adopted by GS13.
In this regime, the host systems are early-type galaxies that feature intermediate-scale disks\footnote{Intermediate-scale disks are 
disks of stars fully embedded in the spheroidal component of their galaxy. 
They are typical of ``disky'' elliptical galaxies (e.g.~NGC 3377), 
but they can also be found in other types of host galaxies.
They can be considered an intermediate class between nuclear disks, with sizes $\sim$$10-100\rm~pc$, %%
and large-scale disks, that encase the bulge and dominate the light at large radii.}. 
Past bulge/disk decompositions failed to correctly identify the extent of such disks and treated them as large-scale disks, 
thus underestimating the bulge luminosity.
The magnitudes that we obtained for such spheroids are on average $\sim$$0.5\rm~mag$ brighter than in GS13. 
These two effects explain the shallower slope that we obtained for the S\'ersic spheroids. \\
The slope that we obtained for S\'ersic spheroids ($1.53 \pm 0.20$) is not consistent with the value of $2.5$ 
expected from $M_{\rm BH} \propto \sigma^5$ and $L_{\rm sph} \propto \sigma^2$. 
In addition, the S\'ersic and core-S\'ersic spheroids appear not to define two distinct $M_{\rm BH} - L_{\rm sph}$ sequences. 
This leads us to investigate substructure in the $M_{\rm BH} - L_{\rm sph}$ diagram for early- and late-type galaxies.
First, we checked that the elliptical and lenticular galaxies, taken separately, have slopes consistent with each other, 
and thus, taken together, they define a single \emph{early-type sequence} in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
We then fit the early-type galaxies with a single log-linear regression, 
and obtained $M_{\rm BH} \propto L_{\rm sph}^{1.00 \pm 0.10}$. 
We did not find any convincing evidence for the change in slope required for consistency 
with the $M_{\rm BH} - \sigma$ and bent $L_{\rm sph} - \sigma$ correlations. 
Because the change in slope should occur at $M_{\rm BH} > 10^{8 \pm 1}\rm~M_\odot$, 
but all the early-type galaxies in our sample have $M_{\rm BH} \gtrsim 10^7\rm~M_\odot$, 
one possible explanation is that we are still not probing enough low black hole masses for this subsample.
An additional possibility is that there is no sharp transition going from $L_{\rm sph} \propto \sigma^2$ at low luminosities 
to $L_{\rm sph} \propto \sigma^5$ at high luminosities. 
Although the knowledge that many ``elliptical'' galaxies actually contain embedded stellar disks dates back  
at least three decades  
\citep{capaccioli1987,carter1987,rixwhite1990,bender1990,scorzabender1990,nieto1991,rixwhite1992,scorzabender1995},
it is mainly thanks to large integral-field-spectrograph surveys of early-type galaxies, such as the ATLAS$^{\rm 3D}$ Project \citep{cappellari2011}, 
that our view has been further advanced 
and it is now commonly accepted that most ``elliptical'' galaxies contain disks. 
Past studies that investigated the $L_{\rm sph} - \sigma$ diagram might have failed to identify and consequently model the disks 
in intermediate-luminosity, early-type galaxies, thus overestimating $L_{\rm sph}$ 
and mistakenly producing a sharp bend in the $L_{\rm sph} - \sigma$ correlation, rather than a continuously curved relation 
(with $L_{\rm sph} \propto \sigma^{3-4}$ at intermediate luminosities). \\
For the bulges of late-type galaxies we obtained $M_{\rm BH} \propto L_{\rm sph}^{2.88 \pm 0.68}$. 
From a cursory inspection of Figure \ref{fig:mbhmagsph}, one might be tempted to doubt the statistical significance of this ``tentative'' \emph{late-type sequence}. 
However, a visual inspection of the plotted data requires one to take into account the error bars 
when judging-by-eye the strength of a correlation. 
Similarly, the Pearson's and Spearman's correlation coefficients are not applicable because 
they do not take into account the error bars on our data.
We have therefore relied on the quantitative regression analysis.

\subsubsection{Pseudo- versus classical bulges}

>>>>

\documentclass{aastex6}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

@@

\title{Erratum: ``Supermassive black holes and their host spheroids \\ 
II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram'' (2016, ApJ, 817, 21)}

@@

\keywords{}

\section*{}

The header of column 5 in Table 4 is incorrectly given as $\langle MAG_{\rm sph} \rangle$. 
It should read $\langle M_{\rm *,sph} \rangle$.
